Equalizer\\Programming and User Guide

### 2009-07-28 Developer/ProgrammingGuide/paper.tex
Update Programming Guide
[ ] Will break Win32 build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\title{Equalizer Programming Guide}
\author{Eyescale Software GmbH}

\date{
  \vspace{15cm}
  {\Large\sffamily The official reference for creating parallel,
    scalable OpenGL\texttrademark\ applications with the Equalizer
    parallel rendering framework}\\\vspace{1cm}
  Version 1.5.9 for Equalizer 0.6+svn\\[\medskipamount]
%  \today
}

@@

\pagestyle{empty}
\maketitle
\thispagestyle{empty}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock}

\clearpage

\lstset{language=C++}
\section*{Equalizer 0.6 Programming Guide}
\today
\subsection*{Contributors}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\listoffigures \vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.5.1   & Dec 08, 2008 & Minor updates based on the print version\\
    1.5.2   & Jan 10, 2009 & Added \sref{sCompositor}\\
    1.5.3   & Jan 13, 2009 & Added \sref{sNetwork}\\
    1.5.4   & Feb 09, 2009 & Added \sref{sNetUsage}\\
    1.5.5   & Mar 10, 2009 & Updated to reflect new AGL event dispatch\\
    1.5.6   & Mar 23, 2009 & Documented Layout API\\
    1.5.7   & Apr 17, 2009 & Updated Layout API (multiple observers)\\
    1.5.8   & May 30, 2009 & Added \sref{sEqualizers}, updated \sref{sLoadBalancing} and \sref{sSwapBarrier}\\
    1.5.9.  & Jul 27, 2009 & Major review pass before release\\
    \hline
  \end{tabularx}}
\clearpage

@@

\section{Introduction}

Equalizer provides a framework for the development of parallel OpenGL
applications. Equalizer-based applications can run from a single
shared-memory system with one or multiple graphics cards up to
large-scale graphics clusters. This Programming Guide introduces the
programming interface, often using the Equalizer example \textsf{eqPly}
as a guideline.

Equalizer is used for scientific simulation visualization at various
research institutes, for creating electronic art exhibitions as well as
in commercial 3D design applications.

Equalizer is the next step in the evolution of generic parallel programming
interfaces for OpenGL-based visualization applications. Existing
solutions, such as OpenGL Multipipe SDK, Cavelib and VRJuggler,
implement a subset of concepts similar to Equalizer. In other areas,
e.g., tracking device support, they provide more functionality.

@@

If you have any question regarding Equalizer programming, this
programming guide, or other specific problems you encountered, please
direct them to the \textsf{eq-dev} mailing
list\footnote{\link{http://www.equalizergraphics.com/lists.html}}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\label{fStereoCmp}\small A Stereo compound}}
\end{wrapfigure}
Due to the frame consistency between the eye views, this modes scales
very well. The IO requirements for pixel transfer are small and
constant.

@@

Eye compounds in Equalizer are configured using the \textsf{eye}
parameter, limiting the child to render the \textsf{[ LEFT ]} or
\textsf{[ RIGHT ]} eye. Each child compound uses an output frame, which
is connected to an input frame on the destination channel. The
destination channel can also be used to render an eye pass, in which case
it renders in the correct stereo buffer and no output frame is needed.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-1ex}\includegraphics[width=.618\textwidth]{images/Pixel.pdf}
  {\caption{\label{fPixel}\small A pixel compound}}
  \vspace{-6ex}
\end{wrapfigure}
Pixel compounds are similar to 2D compounds. The frusta of the
source rendering units are modified so that each unit renders an evenly
distributed subset of pixels, as shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

@@

OpenGL functionality influenced by the raster position will not work
correctly with pixel compounds, or needs at least special
attention. Among them are: lines, points, sprites, glDrawPixels,
glBitmap, glPolygonStipple. The application can query the current pixel
parameters at run-time to adjust the rendering accordingly.

@@

Equalizer are described in more detail in \sref{sEqualizers}.

@@

@@

@@

\begin{wrapfigure}{r}{.382\textwidth}
%\begin{figure}[h!t]
  \vspace{-2ex}
  \subfigure[]{\label{fObjectDist1}
    \includegraphics[width=.382\textwidth]{images/umlObjectSubclass}}\hfil
  \subfigure[]{\label{fObjectDist2}
    \includegraphics[width=.382\textwidth]{images/umlObjectProxy}}\hfil
  \subfigure[]{\label{fObjectDist3}
    \includegraphics[width=.382\textwidth]{images/umlObjectMI}}%
  {\caption{\label{fObjectDist}Object distribution using subclassing,
      proxies or multiple inheritance}}
  \vspace{-4ex}
\end{wrapfigure}
Distributed objects are addressed using session-unique identifiers,
because pointers to other objects cannot be distributed directly; they
have no meaning on remote nodes. The session used to register
distributed objects is normally the \textsf{eq::Config}, which is
subclassed from \textsf{eq::net::Session}.

@@

For head-mounted displays (HMD), the tracking information is used to
move the frusta with the observer, as shown in \fig{fHMF}. This results
in different projections compared to normal tracking with fixed
projection screens.

>>>>

\title{Equalizer\\Programming and User Guide}
\author{Eyescale Software GmbH}

\date{
  \vspace{14cm}
\begin{textblock}{}(0,4.5)
  \hspace{-1cm}\includegraphics[width=22cm]{images/teaser.pdf}
\end{textblock} {\Large The official reference for developing and
  deploying parallel, scalable OpenGL\texttrademark\ applications based on
  the Equalizer parallel rendering framework}\\\vspace{1cm}
Version 1.5.10 for Equalizer 0.6+svn\\[\medskipamount]
%  \today
}

@@

\pagestyle{empty}
\maketitle
\thispagestyle{empty}

%\clearpage

\lstset{language=C++}
\section*{Equalizer 0.6 Programming and User Guide}
\today
\subsection*{Contributors}

@@

\clearpage
\pagenumbering{roman}
\pagestyle{headings}
\tableofcontents
\listoffigures \vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Revision for Equalizer 0.5\\
    1.4     & Nov 25, 2008 & Revision for Equalizer 0.6\\
    1.5.1   & Dec 08, 2008 & Minor updates based on the print version\\
    1.5.2   & Jan 10, 2009 & Added \sref{sCompositor}\\
    1.5.3   & Jan 13, 2009 & Added \sref{sNetwork}\\
    1.5.4   & Feb 09, 2009 & Added \sref{sNetUsage}\\
    1.5.5   & Mar 10, 2009 & Updated to reflect new AGL event dispatch\\
    1.5.6   & Mar 23, 2009 & Documented Layout API\\
    1.5.7   & Apr 17, 2009 & Updated Layout API (multiple observers)\\
    1.5.8   & May 30, 2009 & Added \sref{sEqualizers}, updated \sref{sLoadBalancing} and \sref{sSwapBarrier}\\
    1.5.9   & Jul 27, 2009 & Major review pass before release\\
    1.5.10  & Jul 28, 2009 & Split in User and Programming Guide, added \sref{sNamespaces}\\
    \hline
  \end{tabularx}}
\clearpage

@@

\part{User Guide}
\section{Introduction}

Equalizer is the standard middleware for the development and deployment
of parallel OpenGL applications. It enables applications to benefit from
multiple graphics cards, processors and computers to scale rendering
performance, visual quality and display size. An Equalizer-based
application runs unmodified on any visualization system, from a simple
workstation to large scale graphics clusters, multi-GPU workstations and
Virtual Reality installations.

This User and Programming Guide introduces parallel rendering concepts,
the configuration of Equalizer-based applications and programming using
the Equalizer parallel rendering framework.

Equalizer is the next step in the evolution of middleware for scalable
OpenGL applications. Existing solutions, such as OpenGL Multipipe SDK,
Cavelib and VRJuggler, implement a subset of Equalizer. In other areas,
e.g., tracking device support, some of them provide more functionality.

@@

If you have any question regarding Equalizer programming, this guide, or
other specific problems you encountered, please direct them to the
\textsf{eq-dev} mailing
list\footnote{\link{http://www.equalizergraphics.com/lists.html}}.

@@

Due to the frame consistency between the eye views, this modes scales
very well. The IO requirements for pixel transfer are small and
constant.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/EYE.pdf}
  {\caption{\label{fStereoCmp}\small A stereo compound}}
\end{wrapfigure}
Eye compounds in Equalizer are configured using the \textsf{eye}
parameter, limiting the child to render the \textsf{[ LEFT ]} or
\textsf{[ RIGHT ]} eye. Each child compound uses an output frame, which
is connected to an input frame on the destination channel. The
destination channel can also be used to render an eye pass, in which case
it renders in the correct stereo buffer and no output frame is needed.

@@

Pixel compounds are similar to 2D compounds. The frusta of the
source rendering units are modified so that each unit renders an evenly
distributed subset of pixels, as shown in \fig{fPixel}\footnote{3D model courtesy of AVS, USA.}

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/Pixel.pdf}
  {\caption{\label{fPixel}\small A pixel compound}}
\end{wrapfigure}
OpenGL functionality influenced by the raster position will not work
correctly with pixel compounds, or needs at least special
attention. Among them are: lines, points, sprites, glDrawPixels,
glBitmap, glPolygonStipple. The application can query the current pixel
parameters at run-time to adjust the rendering accordingly.

@@

Equalizers are described in more detail in \sref{sEqualizers}.

@@

\part{Programming Guide}

@@

\subsection{\label{sNamespaces}Namespaces}

Application developers are exposed to the following namespaces:

\begin{description}
\item[eq::base] The eq::base namespace provides C++ classes to abstract
  the underlying operating system and implements common helper
  functionality. One example is the \textsf{eq::base::Clock} class
  providing a high-resolution timer. Classes in this namespace are well
  documented on the API documentation on the Equalizer website, and are
  not subject of this Programming Guide.
\item[eq::net] The Equalizer network layer provides basic functionality
  for network communication, such as \textsf{Connection} and
  \textsf{ConnectionSet}, as well as higher-level functionality such as
  \textsf{Node}, \textsf{Session} and \textsf{Object}. Please refer to
  \sref{sNetwork} for an introduction into the network layer, and to
  \sref{sNetObject} for distributed objects.
\item[eq] The core namespace for the Equalizer client library. The class
  hierarchy in this namespace closely models the configuration file
  format. The classes in the \textsf{eq} namespace are the main subject
  of this Programming Guide.
\end{description}

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/namespaces}
  {\caption{\label{fNamespaces}Namespace layering for eqPly and eqServer}}
\end{wrapfigure}
Examples shipped with Equalizer are implemented in their own namespaces,
e.g., \textsf{eqPly} or \textsf{eVolve}. They rely mostly on subclassing
classes from the \textsf{eq} namespace, with the occasional usage of
functionality from the \textsf{eq::net} and \textsf{eq::base} namespace.

The functionality of the Equalizer server is implemented in the
\textsf{eq::server} namespace. This namespace does not yet expose an API
intended to be used by developers.

\fig{fNamespaces} shows the namespace layering for \textsf{eqPly} and
\textsf{eqServer}.

@@

\begin{wrapfigure}{r}{.382\textwidth}
%\begin{figure}[h!t]
  \subfigure[]{\label{fObjectDist1}
    \includegraphics[width=.382\textwidth]{images/umlObjectSubclass}}\hfil
  \subfigure[]{\label{fObjectDist2}
    \includegraphics[width=.382\textwidth]{images/umlObjectProxy}}\hfil
  \subfigure[]{\label{fObjectDist3}
    \includegraphics[width=.382\textwidth]{images/umlObjectMI}}%
  {\caption{\label{fObjectDist}Object distribution using subclassing,
      proxies or multiple inheritance}}
\end{wrapfigure}
Distributed objects are addressed using session-unique identifiers,
because pointers to other objects cannot be distributed directly; they
have no meaning on remote nodes. The session used to register
distributed objects is normally the \textsf{eq::Config}, which is
subclassed from \textsf{eq::net::Session}.

@@

For head-mounted displays (HMD), the tracking information is used to
move the frusta with the observer, as shown in \fig{fHMD}. This results
in different projections compared to normal tracking with fixed
projection screens.
